\thispagestyle{empty}
\chapter*{Acknowledgments}
 
%Supervisors
I would like to thank my first supervisor and institute director Prof. Frank Kirchner for giving me the opportunity to work on this interesting topic at the DFKI Robotics Innovation Center. Then, I would like to thank my second supervisor Dr. Tobias Bruckmann (Research Group Leader Robotics, Chair of Mechatronics, University of Duisburg-Essen) for providing his continuous guidance and constructive feedback during this thesis.   

%Mentors
I would like to express my deepest gratitude to my main mentor and team leader Dr. Shivesh Kumar for his guidance, the numerous inspiring discussions we had throughout the course of this work, and his infectious enthusiasm for everything related to dynamics and many more surrounding topics. Thanks also goes to Dr. Carlos Mastalli (Postdoctoral researcher, University of Edinburgh, United Kingdom) and Dr. Olivier Stasse (Postdoctoral researcher, LAAS-CNRS, France) who supported me with many ideas and critical but constructive questions related to numerical optimization and humanoid robotics, respectively. 
%Special thanks to my work colleagues Vinzenz Bargsten, Heiner Peters and Daniel Harnack for the fruitful collaboration.   

%Friends, Family, monique
Finally, I want to also thank all my friends I met along the way, in the Lower Rhine area, Duisburg/Essen, Dortmund, %M{\"u}nchen, 
Oberpfaffenhofen, Bremen and elsewhere. You have each made this journey an unforgettable experience. Last but not least, I want to thank my family (especially my mother) and my girlfriend Monique (one of the bravest person I ever met) for your supportive stronghold and I am forever grateful to have you in my life. 


\bigskip

\begin{flushright}
{\Large Julian} \\
%Julian \\
\bigskip
September 18, 2020
\end{flushright}










