%-----------------------------------------------------------------------------%
%                                                                             %
%    K A P I T E L   7                                                       %
%                                                                             %
%-----------------------------------------------------------------------------%

\chapter{Conclusion and Outlook}\label{c7}

\section{Thesis Summary}
This master's thesis was motivated by the generation of physically consistent, efficient motion plans for legged robots. 

The premise of the investigation was that whole-body planning leads to more efficient motions than a simple program such as an \gls{IK} solver. Therefore, the core algorithm of the proposed motion planning approach was a recently presented \gls{DDP}-based whole-body \gls{TO}. Building upon this, a generic method for constraining DDP-like solvers was presented to generate dynamically balanced motions. The results were integrated into the recently presented open-source framework Crocoddyl.

Following this, we investigated the \gls{CoP}-based contact stability of the  proposed motion planning approach for a wide range of motions with the biologically inspired RH5 humanoid robot. We were able to demonstrate that the resulting motion plans for both dynamic bipedal walking and various jumping tasks are inherently balanced. Additionally, the analysis of highly-dynamic movements allowed the derivation of useful guidelines for future design iterations of the humanoid robot.

Although the focus of this thesis was on motion planning, we evaluated the feasibility of the generated trajectories with a simple online stabilizer. We demonstrated in a real-time physics simulator that the motion plans can be stabilized by a simple control architecture solely based on joint-space position control. Furthermore, it could be shown that for real-world experiments, a control in the task space is indispensable to compensate for deviations between the model and reality. 

The final result of this thesis is an efficient motion planning approach that produces inherently balanced motions. This algorithm efficiently generates highly-dynamic movements with flight-phases for various legged systems. 

\section{Future Directions}
%DDP is great
We see large potential in using \gls{DDP}-based whole-body \gls{TO} to generate motions for legged systems. Motion planning based on numerical optimization clearly reduces the number of hand-crafted components and allows the specification of high-level tasks directly in the operational space of the robot. This is likely to become even more important as legged robots tackle more difficult terrains that require higher dynamic motions.
  
%Algorithmic Improvements
From an algorithmic perspective, the formulation can still be improved in a number of ways. The goal of the algorithm is to efficiently and accurately generate physically consistent motions. Promising ways of simultaneously improving both measures are seen by directly embedding inequality constraints as strict bounds inside the \gls{DDP} algorithm, instead of forcing them by penalization.

%Control Approaches: MPC + Online Stabilization
From a control perspective, two successive steps are of particular interest. First, working on an improved online stabilization is a worthwhile undertaking. As discussed in the last chapter, operational space control is inevitable to compensate for modeling errors directly in the task space. Following up, it would be interesting to embed the motion planning inside a \gls{MPC} formulation. If the re-planning is quick enough, the robot will be enabled to act more robustly to unpredicted situations. 

%Long-term vision
This direction for future research, namely improving the motion planning algorithm and embedding it in an \gls{MPC} formulation on a real system, seems promising. Intelligently combining optimization-based planning and control may be the key to robots interacting with both the environment and humans in a more natural, dynamic and autonomous way. With some work in this direction, we may soon find legged robots crossing our daily paths or intuitively collaborating with us when assembling infrastructures and exploring foreign planets. 